---
title: "STA 135 Project"
output: html_document
date: "2024-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
```

```{r}
data(PimaIndiansDiabetes)
```

```{r}
summary(PimaIndiansDiabetes)
```
Pregnant: Number of times pregnant
Glucose : Plasma glucose concentration (glucose tolerance test)
pressure	Diastolic blood pressure (mm Hg) Note: Already kind of weird they aren't doing the S
triceps:	Triceps skin fold thickness (mm)
insulin:	2-Hour serum insulin (mu U/ml)
mass:	Body mass index
pedigree:	Diabetes pedigree function
age:	Age (years)
diabetes:	Class variable (test for diabetes) Note: We are going to try to predict this

However looking at this summary we can see that there is some variables that shouldn't contain a "0" value and should instead contain an NA. This is further supported in the help file of this dataset. 
```{r}
table(PimaIndiansDiabetes$pressure)
```
For example, we can see that there is 35 people that have a blood pressure value of 0 mm Hg. Since this is an impossible reading to receive on an alive human, we will remove these 0 values and replace them with a missing value(NA).

After we have converted all of those false values into NA's we will simply replace those values using the mean of that specific feature/column.
```{r}
cleanData <- function(column) {
  medianValue <- median(column[column != 0], na.rm = TRUE)
  column[column == 0] <- medianValue
  return(column)
}

PimaIndiansDiabetesClean <- PimaIndiansDiabetes


PimaIndiansDiabetesClean$glucose <- cleanData(PimaIndiansDiabetes$glucose)
PimaIndiansDiabetesClean$pressure <- cleanData(PimaIndiansDiabetes$pressure)
PimaIndiansDiabetesClean$triceps <- cleanData(PimaIndiansDiabetes$triceps)
PimaIndiansDiabetesClean$insulin <- cleanData(PimaIndiansDiabetes$insulin)
PimaIndiansDiabetesClean$mass <- cleanData(PimaIndiansDiabetes$mass)
```

```{r}
summary(PimaIndiansDiabetesClean)
```

```{r}
table(PimaIndiansDiabetesClean$pressure)
```

